---
title: "alarm_description"
output: html_document
date: '2022-08-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset description

**ALARM monitoring system (synthetic) data set**


### Description

The **ALARM** ("A Logical Alarm Reduction Mechanism") is a Bayesian network designed to provide an alarm message system for patient monitoring.

### Usage in R

**data(alarm)**

### Format

The alarm data set contains the following 37 variables:

    CVP (central venous pressure): a three-level factor with levels LOW, NORMAL and HIGH.

    PCWP (pulmonary capillary wedge pressure): a three-level factor with levels LOW, NORMAL and HIGH.

    HIST (history): a two-level factor with levels TRUE and FALSE.

    TPR (total peripheral resistance): a three-level factor with levels LOW, NORMAL and HIGH.

    BP (blood pressure): a three-level factor with levels LOW, NORMAL and HIGH.

    CO (cardiac output): a three-level factor with levels LOW, NORMAL and HIGH.

    HRBP (heart rate / blood pressure): a three-level factor with levels LOW, NORMAL and HIGH.

    HREK (heart rate measured by an EKG monitor): a three-level factor with levels LOW, NORMAL and HIGH.

    HRSA (heart rate / oxygen saturation): a three-level factor with levels LOW, NORMAL and HIGH.

    PAP (pulmonary artery pressure): a three-level factor with levels LOW, NORMAL and HIGH.

    SAO2 (arterial oxygen saturation): a three-level factor with levels LOW, NORMAL and HIGH.

    FIO2 (fraction of inspired oxygen): a two-level factor with levels LOW and NORMAL.

    PRSS (breathing pressure): a four-level factor with levels ZERO, LOW, NORMAL and HIGH.

    ECO2 (expelled CO2): a four-level factor with levels ZERO, LOW, NORMAL and HIGH.

    MINV (minimum volume): a four-level factor with levels ZERO, LOW, NORMAL and HIGH.

    MVS (minimum volume set): a three-level factor with levels LOW, NORMAL and HIGH.

    HYP (hypovolemia): a two-level factor with levels TRUE and FALSE.

    LVF (left ventricular failure): a two-level factor with levels TRUE and FALSE.

    APL (anaphylaxis): a two-level factor with levels TRUE and FALSE.

    ANES (insufficient anesthesia/analgesia): a two-level factor with levels TRUE and FALSE.

    PMB (pulmonary embolus): a two-level factor with levels TRUE and FALSE.

    INT (intubation): a three-level factor with levels NORMAL, ESOPHAGEAL and ONESIDED.

    KINK (kinked tube): a two-level factor with levels TRUE and FALSE.

    DISC (disconnection): a two-level factor with levels TRUE and FALSE.

    LVV (left ventricular end-diastolic volume): a three-level factor with levels LOW, NORMAL and HIGH.

    STKV (stroke volume): a three-level factor with levels LOW, NORMAL and HIGH.

    CCHL (catecholamine): a two-level factor with levels NORMAL and HIGH.

    ERLO (error low output): a two-level factor with levels TRUE and FALSE.

    HR (heart rate): a three-level factor with levels LOW, NORMAL and HIGH.

    ERCA (electrocauter): a two-level factor with levels TRUE and FALSE.

    SHNT (shunt): a two-level factor with levels NORMAL and HIGH.

    PVS (pulmonary venous oxygen saturation): a three-level factor with levels LOW, NORMAL and HIGH.

    ACO2 (arterial CO2): a three-level factor with levels LOW, NORMAL and HIGH.

    VALV (pulmonary alveoli ventilation): a four-level factor with levels ZERO, LOW, NORMAL and HIGH.

    VLNG (lung ventilation): a four-level factor with levels ZERO, LOW, NORMAL and HIGH.

    VTUB (ventilation tube): a four-level factor with levels ZERO, LOW, NORMAL and HIGH.

    VMCH (ventilation machine): a four-level factor with levels ZERO, LOW, NORMAL and HIGH.
    
### Source

Beinlich I, Suermondt HJ, Chavez RM, Cooper GF (1989). "The ALARM Monitoring System: A Case Study with Two Probabilistic Inference Techniques for Belief Networks". Proceedings of the 2nd European Conference on Artificial Intelligence in Medicine, 247–256.

### Examples


```{r}
library(bnlearn)
data(alarm)
# create and plot the network structure.
modelstring = paste0("[HIST|LVF][CVP|LVV][PCWP|LVV][HYP][LVV|HYP:LVF][LVF]",
  "[STKV|HYP:LVF][ERLO][HRBP|ERLO:HR][HREK|ERCA:HR][ERCA][HRSA|ERCA:HR][ANES]",
  "[APL][TPR|APL][ECO2|ACO2:VLNG][KINK][MINV|INT:VLNG][FIO2][PVS|FIO2:VALV]",
  "[SAO2|PVS:SHNT][PAP|PMB][PMB][SHNT|INT:PMB][INT][PRSS|INT:KINK:VTUB][DISC]",
  "[MVS][VMCH|MVS][VTUB|DISC:VMCH][VLNG|INT:KINK:VTUB][VALV|INT:VLNG]",
  "[ACO2|VALV][CCHL|ACO2:ANES:SAO2:TPR][HR|CCHL][CO|HR:STKV][BP|CO:TPR]")
dag = model2network(modelstring)
```





