---
title: "lizards_description"
output: html_document
date: '2022-08-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lizards' perching behaviour data set


## Description

Real-world data set about the perching behaviour of two species of lizards in the South Bimini island, from Shoener (1968).

## Usage
```{r}
library(bnlearn)
data(lizards)
```

## Format

The lizards data set contains the following variables:

    Species (the species of the lizard): a two-level factor with levels Sagrei and Distichus.

    Height (perch height): a two-level factor with levels high (greater than 4.75 feet) and low (lesser or equal to 4.75 feet).

    Diameter (perch diameter): a two-level factor with levels narrow (greater than 4 inches) and wide (lesser or equal to 4 inches).

## Source

Edwards DI (2000). Introduction to Graphical Modelling. Springer, 2nd edition.

Fienberg SE (1980). The Analysis of Cross-Classified Categorical Data. Springer, 2nd edition.

Schoener TW (1968). "The Anolis Lizards of Bimini: Resource Partitioning in a Complex Fauna". Ecology, 49(4):704–726.

## Examples

```{r}
# load the data.
data(lizards)
# create and plot the network structure.
dag = model2network("[Species][Diameter|Species][Height|Species]")
## Not run: graphviz.plot(dag, shape = "ellipse")

# This data set is useful as it offers nominal values for
# the conditional mutual information and X^2 tests.
ci.test("Height", "Diameter", "Species", test = "mi", data = lizards)
ci.test("Height", "Diameter", "Species", test = "x2", data = lizards)
```
