---
title: "coronary_description"
output: html_document
date: '2022-08-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Coronary heart disease data set

## Description

Probable risk factors for coronary thrombosis, comprising data from 1841 men.

## Usage
```{r}
library(bnlearn)
data(coronary)
```

## Format

The coronary data set contains the following 6 variables:

    Smoking (smoking): a two-level factor with levels no and yes.

    M. Work (strenuous mental work): a two-level factor with levels no and yes.

    P. Work (strenuous physical work): a two-level factor with levels no and yes.

    Pressure (systolic blood pressure): a two-level factor with levels <140 and >140.

    Proteins (ratio of beta and alpha lipoproteins): a two-level factor with levels <3 and >3.

    Family (family anamnesis of coronary heart disease): a two-level factor with levels neg and pos.

## Source

Edwards DI (2000). Introduction to Graphical Modelling. Springer, 2nd edition.

Reinis Z, Pokorny J, Basika V, Tiserova J, Gorican K, Horakova D, Stuchlikova E, Havranek T, Hrabovsky F (1981). "Prognostic Significance of the Risk Profile in the Prevention of Coronary Heart Disease". Bratisl Lek Listy, 76:137–150. Published on Bratislava Medical Journal, in Czech.

Whittaker J (1990). Graphical Models in Applied Multivariate Statistics. Wiley.
Examples


```{r}
# This is the undirected graphical model from Whittaker (1990).
data(coronary)
ug = empty.graph(names(coronary))
arcs(ug, check.cycles = FALSE) = matrix(
  c("Family", "M. Work", "M. Work", "Family",
    "M. Work", "P. Work", "P. Work", "M. Work",
    "M. Work", "Proteins", "Proteins", "M. Work",
    "M. Work", "Smoking", "Smoking", "M. Work",
    "P. Work", "Smoking", "Smoking", "P. Work",
    "P. Work", "Proteins", "Proteins", "P. Work",
    "Smoking", "Proteins", "Proteins", "Smoking",
    "Smoking", "Pressure", "Pressure", "Smoking",
    "Pressure", "Proteins", "Proteins", "Pressure"),
  ncol = 2, byrow = TRUE,
  dimnames = list(c(), c("from", "to")))
## Not run: graphviz.plot(ug, shape = "ellipse")
```